Improvement of Thermostated Molecular Dynamics via Backward Error Analysis

### 2018-06-08 rigid_bodies.tex
Changes in NVE results

\subsection{Adjustment of the NVT Dynamics}
\label{sec:refined_method}

@@

\section{Application to Water and Discussion}
\label{sec:numerical_results}

In the following, we analyze the performance of the NVT scheme proposed in  this paper and compare the results with those obtained employing the integrator due to \citeauthor{Martyna_1996} \cite{Martyna_1996} as well as the method of \citeauthor{Kamberaj_2005} \cite{Kamberaj_2005}

The system under study corresponds to 903 TIP3P\cite{Jorgensen_1983} water molecules at a density of 970 kg/m$^3$.
The 12-6 Lennard-Jones and damped Coulombic interactions were truncated at  10 {\AA} and smoothed via a switching function over the range of 9.5 {\AA} to  10 {\AA} as done in Ref.~\citenum{Silveira_2017}. 
The damping of Coulombic interactions was done by a factor erfc($\alpha r$), with $\alpha$ = 0.29 {\AA}$^{-1}$.
In addition, all the simulations were carried out considering $t_d = 100$ fs.

Before presenting our main results, we first examine the departure of the \textit{refined} Hamiltonian from the specified one, as this is the most direct way to determine the importance of the discretization erros \cite{Engle_2005}, which is essential for discussing the NVT case.
In Fig.~\ref{fig:nve} we show the time evolution of both $\refined{\Ham{}}$ (dotted lines) and $\Ham{}$ (solid lines) for the water system undergoing NVE dynamics with different time-step sizes ($\timestep$).
For this, we used the numerical scheme of Eq.~\ref{eq:splitting NVE} with the unsplit solution of rotation.
As expected, $\refined{\Ham{}}$ is very well conserved and close to $\Ham{}$ for $\timestep$ = 1 fs, but fairly rapidly separates from $\Ham{}$ when increasing $\timestep$. 
It should be remarked that $\refined{\Ham{}}$ corresponds to a simple approximation and, most important, it is computationally very cheap.
As a result of such approximation, we have larger fluctuations in $\refined{\Ham{}}$ for $\timestep$ = 3 fs and $\timestep$ = 4 fs (add in conclusions that we should use high-order approximations).

>>>>

\subsection{Refinement of the NVT Dynamics}
\label{sec:refined_method}

@@

\section{Application to the Simulation of Liquid Water}
\label{sec:numerical_results}

In this section, we analyze the performance of our proposed NVT-MD method and compare its results to those obtained by employing both the method of \citeauthor{Kamberaj_2005} \cite{Kamberaj_2005} and the method of \citeauthor{Martyna_1996} \cite{Martyna_1996} with our described adaptations.

The system under study has periodic boundary conditions and contains 903 TIP3P \cite{Jorgensen_1983} water molecules at a density of 970~kg/m\textsuperscript{3}.
The 12-6 Lennard-Jones and damped Coulombic interactions were truncated at 10 {\AA} and smoothed by a switching function over the range from 9.5 \AA to 10 \AA, such as done in Ref.~\citenum{Silveira_2017}. 
The damping of Coulombic interactions was done by a factor $\text{erfc}(-\alpha r)$, with $\alpha = 0.29~\text{\AA}^{-1}$.
In addition, all thermostated simulations were carried out considering a time-scale constant $t_d = 100~\text{fs}$.

We start by examining the departure of the \textit{refined} Hamiltonian from the specified one in NVE simulations.
This is the most direct way of determining the importance of the discretization errors \cite{Engle_2005}, which is essential for discussing the NVT case.
In Fig.~\ref{fig:nve} we show the time evolution of both $\refined{\Ham{}}$ (dotted lines) and $\Ham{}$ (solid lines) for the water system using different time-step sizes ($\timestep$).
For this, we used the numerical scheme of Eq.~\eqref{eq:splitting NVE} with the unsplit solution of Ref.~\citenum{Silveira_2017} applied for the rotations.
The four-point differentiation formulas presented in Sec.~\ref{sec:refined_method} were employed for computing the refined kinetic energy.
As expected, $\refined{\Ham{}}$ is very well conserved and close to $\Ham{}$ for $\timestep = 1~\text{fs}$, but rapidly departs from $\Ham{}$ when $\timestep$ increases.
It is worth remarking that $\refined{\Ham{}}$ corresponds to a simple and, most importantly, computationally inexpensive approximation of the truly conserved shadow Hamiltonian.
As a result, we observe increasing fluctuations in $\refined{\Ham{}}$ for $\timestep = 3~\text{fs}$ and $\timestep = 4~\text{fs}$.
This is an indication that
1) the employed four-point formulas do not provide derivatives with the required accuracy or
2) the high-order terms of the Baker-Campbell-Hausdorff series (see Appendix \ref{sec:rigid body shadow hamiltonian}) would have a non-negligible impact on the shadow Hamiltonian.
